---
title: "Are Snakes Left-handed?"
author: "Troy Edwards"
date: "2023-02-16"
output: pdf_document
fontsize: 12pt
documentclass: article
geometry:
  margin=1in
---

```{r setup, include=FALSE}
setwd("C:\\Users\\tntje\\work\\school\\applied-stats\\s2\\assignment04")

library(ggplot2)
library(GGally)
library(dplyr)

knitr::opts_chunk$set(warning = FALSE, message = FALSE) 

snake_data <- read.csv("data.csv")
snake_data$Cat.Pair <- paste0(substr(snake_data$Age, 1, 1), substr(snake_data$Gender, 1, 1))
```

# Question 1

*Do these creatures in general exhibit asymmetric coiling behavior? (Include a
graphical display and an inference test to answer this question.)*

## Graph

```{r q1g, echo=FALSE}
ggplot(snake_data, aes(Lat.Index)) +
  geom_histogram(aes(fill = ifelse(Lat.Index > 0.5, "Left", "Right")), bins = 24) +
  scale_x_continuous(limits = c(0.35, 0.65)) +
  scale_fill_manual(
    name = "Coiling direction tendency",
    values = setNames(c('#ff8888ff', '#8888ffff'), c("Left", "Right"))
  ) +
  labs(
    title = "Distribution of laterality index",
    x = "Laterality index",
    y = "Count"
  )
```

## Inference Test

```{r q1t, include=FALSE}
mean(snake_data$Lat.Index)
sd(snake_data$Lat.Index)
t.test(snake_data$Lat.Index, mu = 0.5, alternative = "two.sided")
```

\begin{center}
Let $\mu =$ the true mean laterality index of cottonmouth snakes \\
$H_0: \mu = 0.5$ \\
$H_a: \mu \neq 0.5$ \\
$\alpha = 0.05$ \\
Test: 1-sample mean t-test \\
\end{center}

### Checks

Normal: Central limit theorem, $n = 36 \geq 30$  
Random: Reasonable to assume that SRS was used  
10%: Reasonable to assume that $n \leq 0.1N$

### Calculations

\begin{center}
$t = \frac{\bar{x} - \mu}{\sqrt{\frac{s^2}{n}}} = \frac{0.5184 - 0.5}{\sqrt{\frac{0.0566^2}{36}}} = 1.9468$ \\
$p = P(t > 1.9468) = 0.0596$ (with $df = 35$)
\end{center}

### Conclusion

Since our $p$-value (0.0596) is greater than $\alpha$ (0.05), we fail to reject
our null hypothesis that cottonmouth snakes generally exhibit symmetric
coiling behavior. Therefore, we do not have sufficient evidence to support the
alternative hypothesis that cottonmouth snakes exhibit asymmetric coiling
behavior. 

\pagebreak

# Question 2

*Which individual category (or combination of categories) of snake best
exhibits/predicts asymmetric coiling behavior? (Explore different linear and
multiple regression models. Chose the best fit. Include supporting
documentation as to why you chose the model you did. No need to run an
inference test.)*

## Models

### Gender

```{r q2reg1, echo=TRUE}
regres1 <- lm(Lat.Index ~ factor(Gender), data = snake_data)
summary(regres1)
```

### Age

```{r q2reg2, echo=TRUE}
regres2 <- lm(Lat.Index ~ factor(Age), data = snake_data)
summary(regres2)
```

### Gender:Age

```{r q2reg3, echo=TRUE}
regres3 <- lm(Lat.Index ~ factor(Gender) : factor(Age), data = snake_data)
summary(regres3)
```

## Conclusion

It appears that a combination of age and gender is the best way to predict a
snake's asymmetric coiling behavior. The model using both age and gender had an
$r^2$ value of 0.4154 and an $r^2(adj)$ value of 0.3605. The model using only
gender and the one using only age had $r^2$ values of 0.105 and 0.2405,
respectively, and $r^2(adj)$ values of 0.07871 and 0.2181, respectively. 

# Question 3

*Do adults differ from juveniles in their asymmetric coiling behavior? Also, do
males differ from females in their asymmetric coiling behavior? (Two separate
inference tests are needed to answer this prompt.)*

## Adults vs. Juveniles

\begin{center}
Let $\mu_a$ represent the true mean laterality index of adult cottonmouth snakes \\
Let $\mu_j$ represent the true mean laterality index of juvenile cottonmouth snakes \\
$H_0: \mu_a = \mu_j$ \\
$H_a: \mu_a \neq \mu_j$ \\
$\alpha = 0.05$ \\
Test: 2-sample difference of means t-test \\
\end{center}

### Checks

Normal: $n_a = 18 < 30$ and $n_j = 18 < 30$, normal probability plots required

```{r q3qqp1, echo=FALSE, fig.height=unit(2.25, "in"), fig.width=unit(6.5, "in")}
ggplot(data = snake_data[snake_data$Age == "Adult",], aes(sample = Lat.Index)) +
  geom_qq() +
  labs(
    x = "Theoretical",
    y = "Sample",
    title = "Normal Q-Q plot of laterality index of adult cottonmouth snakes"
  )
```

```{r q3qqp2, echo=FALSE, fig.height=unit(2.25, "in"), fig.width=unit(6.5, "in")}
ggplot(data = snake_data[snake_data$Age == "Juvenile",], aes(sample = Lat.Index)) +
  geom_qq() +
  labs(
    x = "Theoretical",
    y = "Sample",
    title = "Normal Q-Q plot of laterality index of juvenile cottonmouth snakes"
  )
```

Since both Q-Q plots appear relatively linear, we will proceed with the test.  
Random: reasonable to assume SRS was used  
10%: reasonable to assume that $n_a \leq 0.1N_a$ and $n_j \leq 0.1N_j$

### Calculations

```{r q3t1, include=FALSE}
sd(snake_data[snake_data$Age == "Adult",]$Lat.Index)
sd(snake_data[snake_data$Age == "Juvenile",]$Lat.Index)
t.test(Lat.Index ~ Age, data = snake_data)
```

\begin{center}
$t = \frac{(\bar{x_a} - \bar{x_j}) - (\mu_a - \mu_j)}{\sqrt{\frac{s_a^2}{n_a} + \frac{s_j^2}{n_j}}}
= \frac{(0.5457 - 0.4910) - (0)}{\sqrt{\frac{0.0589^2}{18} + \frac{0.0392^2}{18}}} = 3.2808$ \\
$p = P(t > 3.2808) = 0.002657$ with $df = 29.568$
\end{center}

### Conclusion

Since our $p$-value (0.002657) is less than $\alpha$ (0.05), we reject the null
hypothesis that the true mean laterality indices of adult and juvenile
cottonmouth snakes are equal. Therefore, we have sufficient evidence to
support the alternative hypothesis that adults and juveniles have different
mean laterality indices. 

## Males vs. Females

\begin{center}
Let $\mu_m$ represent the true mean laterality index of male cottonmouth snakes \\
Let $\mu_f$ represent the true mean laterality index of female cottonmouth snakes \\
$H_0: \mu_m = \mu_f$
$H_a: \mu_m \neq \mu_f$
$\alpha = 0.05$
Test: 2-sample difference of means t-test
\end{center}

### Checks

Normal: $n_m = 18 < 30$ and $n_f = 18 < 30$, normal probability plots required

```{r q3qqp3, echo=FALSE, fig.height=unit(2.25, "in"), fig.width=unit(6.5, "in")}
ggplot(data = snake_data[snake_data$Gender == "Male",], aes(sample = Lat.Index)) +
  geom_qq() +
  labs(
    x = "Theoretical",
    y = "Sample", 
    title = "Normal Q-Q plot of laterality index of male cottonmouth snakes"
  )
```

```{r q3qqp4, echo=FALSE, fig.height=unit(2.25, "in"), fig.width=unit(6.5, "in")}
ggplot(data = snake_data[snake_data$Gender == "Female",], aes(sample = Lat.Index)) +
  geom_qq() +
  labs(
    x = "Theoretical",
    y = "Sample", 
    title = "Normal Q-Q plot of laterality index of female cottonmouth snakes"
  )
```
Since both Q-Q plots appear relatively linear, we will proceed with the test.  
Random: reasonable to assume SRS was used  
10%: reasonable to assume that $n_m \leq 0.1N_m$ and $n_f \leq 0.1N_f$

### Calculations

```{r q3t2, include=FALSE}
sd(snake_data[snake_data$Gender == "Male",]$Lat.Index)
sd(snake_data[snake_data$Gender == "Female",]$Lat.Index)
t.test(Lat.Index ~ Gender, data = snake_data)
```

\begin{center}
$t = \frac{(\bar{x_m} - \bar{x_f}) - (\mu_m - \mu_f)}{\sqrt{\frac{s_m^2}{n_m} + \frac{s_f^2}{n_f}}}
= \frac{(0.5003 - 0.534) - (0)}{\sqrt{\frac{0.0456^2}{18} + \frac{0.618^2}{18}}} = 1.9975$ \\
$p = P(t > 1.9975) = 0.05453$ w/ $df = 31.278$
\end{center}

### Conclusion

Since our $p$-value (0.05453) is greater than $\alpha$ (0.05), we fail to
reject the null hypothesis that the true mean laterality indices of male and
female cottonmouth snakes are different. Therefore, we do not have significant
evidence to support the alternative hypothesis that the true mean laterality
indices of male and female cottonmouth snakes are not equal. 

## Conclusion

In conclusion, juveniles differ from adults in their coiling behavior, while
males do not differ from females. 

# Question 4

*Is there a difference in mean coiling laterality index per each categorical
pairing of snake? (AF, JF, AM, JM) (Well, if you have been paying attention
during the new learning from Semester II, this question should be easy for you
to answer. State hypotheses, show all work with your calculations, confirm your
calculations with an ANOVA summary table using R which you will copy and paste
into your narrative/document, and write a conclusion.)*

## Inference test

\begin{center}
Let $\mu_{af}$, $\mu_{jf}$, $\mu_{am}$, and $\mu_{jm}$ represent the true mean
laterality indices of adult female, juvenile female, adult male, and juvenile
male cottonmouth snakes, respectively \\
$H_0: \mu_{af} = \mu_{jf} = \mu_{am} = \mu_{jm}$ \\
$H_a: H_0$ is false (i.e. at least one of the means is different from the
others) \\
$\alpha = 0.05$ \\
Test: One-Way ANOVA
\end{center}

### Checks

All populations approximately normal: 

```{r q4qqp1, echo=FALSE, fig.height=unit(2.25, "in"), fig.width=unit(6.5, "in")}
ggplot(data = snake_data[snake_data$Age == "Adult" & snake_data$Gender == "Female",], aes(sample = Lat.Index)) +
  geom_qq() +
  labs(
    x = "Theoretical",
    y = "Sample",
    title = "Normal Q-Q plot of laterality index for adult female cottonmouth snakes"
  )
```

```{r q4qqp2, echo=FALSE, fig.height=unit(2.25, "in"), fig.width=unit(6.5, "in")}
ggplot(data = snake_data[snake_data$Age == "Juvenile" & snake_data$Gender == "Female",], aes(sample = Lat.Index)) +
  geom_qq() +
  labs(
    x = "Theoretical",
    y = "Sample",
    title = "Normal Q-Q plot of laterality index for juvenile female cottonmouth snakes"
  )
```

```{r q4qqp3, echo=FALSE, fig.height=unit(2.25, "in"), fig.width=unit(6.5, "in")}
ggplot(data = snake_data[snake_data$Age == "Adult" & snake_data$Gender == "Male",], aes(sample = Lat.Index)) +
  geom_qq() +
  labs(
    x = "Theoretical",
    y = "Sample",
    title = "Normal Q-Q plot of laterality index for adult male cottonmouth snakes"
  )
```

```{r q4qqp4, echo=FALSE, fig.height=unit(2.25, "in"), fig.width=unit(6.5, "in")}
ggplot(data = snake_data[snake_data$Age == "Juvenile" & snake_data$Gender == "Male",], aes(sample = Lat.Index)) +
  geom_qq() +
  labs(
    x = "Theoretical",
    y = "Sample",
    title = "Normal Q-Q plot of laterality index for juvenile male cottonmouth snakes"
  )
```

Since the normal Q-Q plots are all approximately linear, we will proceed with the test.  
Standard deviations approximately equal: largest standard deviation is no more than twice the smallest  
$s_{smallest} = 0.014 < 2s_{largest} = 2 \times 0.062$
Check failed, proceeding with caution anyways  
Random: reasonable to assume SRS was used

### Calculations

```{r q4t, include=FALSE}
sd(snake_data[snake_data$Age == "Adult" & snake_data$Gender == "Female",]$Lat.Index)
sd(snake_data[snake_data$Age == "Juvenile" & snake_data$Gender == "Female",]$Lat.Index)
sd(snake_data[snake_data$Age == "Adult" & snake_data$Gender == "Male",]$Lat.Index)
sd(snake_data[snake_data$Age == "Juvenile" & snake_data$Gender == "Male",]$Lat.Index)
aov(Lat.Index ~ Cat.Pair, data = snake_data)
anova(aov(Lat.Index ~ Cat.Pair, data = snake_data))
```

\begin{center}
$SSTr = \sum{n_k(\bar{x}_k - \bar{\bar{x}})^2} = 0.0466$ \\
$SSE = \sum{(n_k - 1)s_k^2} = 0.0655$ \\
$MSTr = \frac{SSTr}{k - 1} = \frac{0.0466}{4 - 1} = 0.0155$ \\
$MSE = \frac{SSE}{N - k} = \frac{0.0655}{36 - 4} = 0.0020$ \\
$F = \frac{MSTr}{MSE} = \frac{0.0155}{0.0020} = 7.5781$ \\
$p = P(F > 7.5781) = 0.0005767$ w/ $df = \frac{3}{32}$
\end{center}

### Conclusion

Since our $p$-value (0.0005767) is less than $\alpha$ (0.05), we reject our
null hypothesis that all of the true means are different. Therefore, we have
enough evidence to support the alternative hypothesis that at least one of the
true means is different from the others. 

## ANOVA table
```{r q4at1, echo=FALSE}
anova(aov(Lat.Index ~ Cat.Pair, data = snake_data))
```