---
title: "Are Snakes Left-handed?"
author: "Troy Edwards"
date: "2023-02-16"
output: pdf_document
fontsize: 12pt
documentclass: article
geometry:
  margin=1in
---

```{r setup, include=FALSE}
setwd("C:\\Users\\tntje\\work\\school\\applied-stats\\s2\\assignment04")

library(ggplot2)
library(GGally)
library(dplyr)

knitr::opts_chunk$set(warning = FALSE, message = FALSE) 

snake_data <- read.csv("data.csv")
```

# 1)

*Do these creatures in general exhibit asymmetric coiling behavior? (Include a
graphical display and an inference test to answer this question.)*

```{r q1, echo=FALSE}
ggplot(snake_data, aes(Lat.Index)) +
  geom_histogram(aes(fill = ifelse(Lat.Index > 0.5, "Left", "Right")), bins = 24) +
  scale_x_continuous(limits = c(0.35, 0.65)) +
  scale_fill_manual(
    name = "Coiling direction tendency",
    values = setNames(c('#ff8888ff', '#8888ffff'), c("Left", "Right"))
  ) +
  labs(
    title = "Distribution of laterality index",
    x = "Laterality index",
    y = "Count"
  )
```

\begin{center}
Let $\mu_L =$ the true mean laterality index for cottonmouth snakes \\
$H_0: \mu_L = 0.5$ (i.e. coiling behavior is symmetric) \\
$H_a: \mu_L \neq 0.5$ (i.e. coiling behavior is asymmetric) \\
$\alpha: 0.05$ \\
Test: one-sample difference of means t-test \\


\end{center}
