---
title: "#9 Logistic Regression: First-Year GPA"
author: "Troy Edwards"
date: "2023-03-26"
output: pdf_document
fontsize: 12pt
documentclass: article
geometry:
  margin=1in
---

```{r setup, include=F}
setwd("C:/Users/tntje/work/school/applied-stats/s2/assignment09")
library(ggplot2)
gpa <- read.csv("FirstYearGPA.csv")
gpa <- na.omit(gpa)
gpa$Male <- factor(gpa$Male)
gpa$FirstGen <- factor(gpa$FirstGen)
gpa$White <- factor(gpa$White)
```

# a)

$$H_0: \mu_0 = \mu_1$$
$$H_a: \mu_0 \neq \mu_1$$
$$\alpha = 0.05$$
$$\text{Test: 2-sample difference in means t-test}$$

Random: assume data was collected using simple random sample  
Normal: $n_0 = 194 \geq 30$, $n_1 = 25 \ngeq 30$, CLT fails, proceeding with caution  
Independent: reasonable to assume that $n \leq 0.1N$  

```{r qa}
t.test(GPA ~ FirstGen, data = gpa)
```

$$t = \frac{\bar{x}_0 - \bar{x}_1 - (\mu_0 - \mu_1)}{\sqrt{\frac{s_0^2}{n_0} + \frac{s_1^2}{n_1}}}$$

Since our p-value (0.02017) is less than $\alpha$ (0.05), we reject our null
hypothesis that the true mean GPA for first-generation college students is
equal to that of non-first-generation college students. Therefore, we have
enough evidence to conclude that the true mean GPAs for these two groups are
not equal. 

# b)

```{r qb}
regres01 <- lm(GPA ~ FirstGen, data = gpa)
summary(regres01)
```

# TODO

# c)

```{r qc}
regres02 <- glm(FirstGen ~ GPA, data = gpa, family = binomial("logit"))
summary(regres02)
```

# \textcolor{red}{TODO}

# d)

```{r qd, include=FALSE}
predict.glm(regres02, newdata = list(GPA = 4.0), type = "response")
predict.glm(regres02, newdata = list(GPA = 1.035642), type = "response")
predict.glm(regres02, newdata = list(GPA = 0.3679345), type = "response")
```

## Part 1

$$ P(FirstGen | GPA = 4.0) = \frac{e^{1.0751 - 1.0381(4.0)}}{1 + e^{1.0751 - 1.0381(4.0)}} = 0.04405638 $$

## Part 2

$$ ln \left( \frac{0.5}{1 - 0.5} \right) = 1.0751 - 1.0381(GPA) $$
$$ 1.0381(GPA) = 1.0751 - ln \left( \frac{0.5}{1 - 0.5} \right) $$
$$ GPA = \frac{1.0751 - ln \left( \frac{0.5}{1 - 0.5} \right)}{1.0381} $$
$$ GPA = \frac{1.0751 - 0}{1.0381} $$
$$ GPA = 1.035642 $$

## Part 3

$$ ln(2) = 1.0751 - 1.0381(GPA) $$
$$ 1.0381(GPA) = 1.0751 - ln(2) $$
$$ GPA = \frac{1.0751 - ln(2)}{1.0381} $$
$$ GPA = 0.3679345 $$

# e)

```{r qe, include=FALSE}
regres03 <- lm(
  GPA ~
    HSGPA +
    SATV +
    SATM +
    Male +
    FirstGen +
    White,
  data = gpa
)
summary(regres03)

regres04 <- lm(
  GPA ~
    HSGPA,
  data = gpa
)
summary(regres04)

regres05 <- lm(
  GPA ~
    HSGPA *
    FirstGen,
  data = gpa
)
summary(regres05)

regres06 <- lm(
  GPA ~
    HSGPA *
    SATV *
    SATM *
    Male *
    FirstGen *
    White,
  data = gpa
)
summary(regres06)

regres07 <- lm(
  GPA ~
    SATV,
  data = gpa
)
summary(regres07)

regres08 <- lm(
  GPA ~
    SATM,
  data = gpa
)
summary(regres08)

regres09 <- lm(
  GPA ~
    Male,
  data = gpa
)
summary(regres09)

regres10 <- lm(
  GPA ~
    FirstGen,
  data = gpa
)
summary(regres10)

regres11 <- lm(
  GPA ~
    White,
  data = gpa
)
summary(regres11)

regres12 <- lm(
  GPA ~
    HSGPA *
    White,
  data = gpa
)
summary(regres12)

regres13 <- lm(
  GPA ~
    HSGPA *
    FirstGen *
    White,
  data = gpa
)
summary(regres13)

regres14 <- lm(
  GPA ~
    HSGPA *
    FirstGen +
    White +
    HSGPA :
    White,
  data = gpa
)
summary(regres14)

regres15 <- lm(
  GPA ~
    HSGPA +
    SATV *
    FirstGen +
    White,
  data = gpa
)
summary(regres15)

regres16 <- lm(
  GPA ~
    HSGPA +
    SATV :
    ifelse(FirstGen == 1, 1, 0) +
    White,
  data = gpa
)
summary(regres16)

regres17 <- lm(
  GPA ~
    HSGPA +
    White,
  data = gpa
)
summary(regres17)
```

\begin{tabular}{ | p{3.5in} | c | c | c | c | }
\hline
Model & $R^2(adj)$ & $S_e$ & $F$ & $p$ \\
\hline
$\widehat{GPA} = 1.17985 + 0.55501(HSGPA)$ & $0.196$ & 0.4174 & $54.15$ & $3.783 \times 10^{-12}$ \\
\hline
$\widehat{GPA} = 1.160826 + 0.569504(HSGPA) - 0.284592(FirstGen) + 0.003677(HSGPA)(FirstGen)$ & $0.2235$ & $0.4102$ & $21.91$ & $2.018 \times 10^{-12}$ \\
\hline
$\widehat{GPA} = 2.0684133 + 0.0016986(SATV)$ & $0.08842$ & $0.4444$ & $22.15$ & $4.499 \times 10^{-6}$ \\
\hline
$\widehat{GPA} = 0.2459 + 0.7593(HSGPA) + 1.3010(White) - 0.2923(HSGPA)(White)$ & $0.2684$ & $0.3981$ & $27.65$ & $2.647 \times 10^{-15}$ \\
\hline
$\widehat{GPA} = 0.99579 + 0.54004(HSGPA) + 0.29842(White)$ & $0.2613$ & $0.4001$ & $39.56$ & $2.303 \times 10^{-15}$ \\
\hline
\end{tabular}

*Note: I did test more models with more predictors but they were too big to put
in the table.*
